Fast approximate Bayesian inference for small-area estimation of HIV indicators using the Naomi model

### 2023-03-29 src/docs_paper/paper.Rmd
Add to-do list for paper

@@

The joint posterior of $(\x, \btheta)$ for an ELGM is given by
\begin{equation*}
  p(\x, \btheta \, | \, \y)
  \propto p(\btheta) |\mathbf{Q}(\btheta)|^{n/2} \exp \left( - \frac{1}{2} \x^\top \mathbf{Q}(\btheta) \x + \sum_{i = 1}^n \log p(y_i \, | \, \x_{\mathcal{J}_i}, \btheta) \right).
\end{equation*}
We consider approximations to the posterior marginals of each latent random variable $x_i$ and hyperparameter $\theta_j$ given by
\begin{align}
  \tilde p(x_i \, | \, \y) \approx p(x_i \, | \, \y) &= \int p(x_i, \btheta \, | \, \y) \text{d} \btheta = \int p(x_i \, | \, \btheta, \y) p(\btheta \, | \, \y) \text{d}\btheta, \quad i \in [N], \label{eq:inla1} \\
  \tilde p(\theta_j \, | \, \y) \approx p(\theta_j \, | \, \y) &= \int p(\btheta \, | \, \y) \text{d}\btheta_{-j} \quad j \in [m]. \label{eq:inla2}
\end{align}
Given the negative unnormalised log posterior $- \log p(\y, \x, \btheta)$, we obtain the above posterior marginal approximations $\{ \tilde p(x_i \, | \, \y) \}_{i = 1}^n$ and $\tilde p(\theta_j \, | \, \y)_{j = 1}^m$ via nested applications of the Laplace approximation and AGHQ, described below.

@@

```{r aghq, fig.cap="The Gauss-Hermite quadrature nodes $\\z \\in \\mathcal{Q}(2, 3)$ for this two dimensional integral with three nodes per dimension are adapted based on the mean and covariance matrix of the target via $\\btheta(\\z) = \\hat \\btheta + \\mathbf{L} \\z$.."}
mu <- c(1, 1.5)
cov <- matrix(c(2, 1, 1, 1), ncol = 2)

>>>>

<!-- Remaining things to add, and or desiderata: -->

<!-- * [ ] Description of the ART attendance component of the model -->
<!-- * [ ] Tying together Naomi model within LGM / ELGM mathematical notation -->
<!-- * [ ] Create DAG representing Naomi model -->
<!-- * [ ] Rewrite section on why Naomi is an ELGM, drawing from slides -->
<!-- * [ ] Write section on Laplace approximation -->
<!-- * [ ] Write section on AGHQ, tying together connection to Laplace for k = 1 -->
<!-- * [ ] Move algorithm to algorithm environment -->

@@

The joint posterior of $(\x, \btheta)$ for an ELGM is given by
\begin{equation*}
  p(\x, \btheta \, | \, \y)
  \propto p(\btheta) |\mathbf{Q}(\btheta)|^{n/2} \exp \left( - \frac{1}{2} \x^\top \mathbf{Q}(\btheta) \x + \sum_{i = 1}^n \log p(y_i \, | \, \x_{\mathcal{J}_i}, \btheta) \right).
\end{equation*}
We consider approximations to the posterior marginals of each latent random variable $x_i$ and hyperparameter $\theta_j$ given by
\begin{align}
  \tilde p(x_i \, | \, \y) \approx p(x_i \, | \, \y) &= \int p(x_i, \btheta \, | \, \y) \text{d} \btheta = \int p(x_i \, | \, \btheta, \y) p(\btheta \, | \, \y) \text{d}\btheta, \quad i \in [N], \label{eq:inla1} \\
  \tilde p(\theta_j \, | \, \y) \approx p(\theta_j \, | \, \y) &= \int p(\btheta \, | \, \y) \text{d}\btheta_{-j} \quad j \in [m]. \label{eq:inla2}
\end{align}
Given the negative unnormalised log posterior $- \log p(\y, \x, \btheta)$, we obtain the above posterior marginal approximations $\{ \tilde p(x_i \, | \, \y) \}_{i = 1}^n$ and $\{\tilde p(\theta_j \, | \, \y)\}_{j = 1}^m$ via nested applications of the Laplace approximation and AGHQ.

@@

```{r aghq, fig.cap="The Gauss-Hermite quadrature nodes $\\z \\in \\mathcal{Q}(2, 3)$ for this two dimensional integral with three nodes per dimension are adapted based on the mean and covariance matrix of the target via $\\btheta(\\z) = \\hat \\btheta + \\mathbf{L} \\z$."}
mu <- c(1, 1.5)
cov <- matrix(c(2, 1, 1, 1), ncol = 2)
